\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,booktabs,microtype}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\newcommand{\Sym}{\operatorname{Sym}}
\newcommand{\cS}{\mathcal S}
\title{Hamming Schemes: Wreath-Product Permutation Formula\\Proof and Exact Verification}
\author{Executable companion note}
\date{17 July 2026}
\begin{document}
\maketitle
\begin{abstract}
We prove and test the generalized Molien formula for the full Hamming graph
automorphism group $S_q\wr S_D=S_q^D\rtimes S_D$ acting on $Q^D$.  The tests
construct every wreath-product element, not merely pure coordinate
permutations, and compare matrix, cycle, and orbit calculations through total
degree three.
\end{abstract}

\section{Permutation Molien and orbit theorem}
For any finite action $G\curvearrowright X$ with vertex-cycle counts
$c_\ell(g)$,
\begin{equation}\label{eq:master}
\cS_{G,X}(\mathbf t)=\frac1{|G|}\sum_{g\in G}
\prod_{i,\ell}(1-t_i^\ell)^{-c_\ell(g)}.
\end{equation}
This follows because an $\ell$-cycle block satisfies
$\det(I-tP_\ell)=1-t^\ell$.  Moreover,
\begin{equation}\label{eq:orbit}
[m_\tau]\cS_{G,X}=\#\left(G\backslash
\prod_j\Sym^{\tau_j}(X)\right),
\end{equation}
since $\Sym^d(\mathbb C[X])$ has its monomial basis indexed by $d$-multisets
of vertices.  This gives a proof route independent of determinant algebra.

\section{Fixed words in the full wreath product}
Write $g=(\sigma_1,\ldots,\sigma_D;\pi)$.  For a fixed power $g^r$, follow a
coordinate around a cycle $C=(i_1,\ldots,i_s)$ of its coordinate permutation.
The successive alphabet relabelings compose to a permutation $h_{C,r}\in S_q$.
A fixed word is determined on $C$ by one symbol fixed by $h_{C,r}$, and
different coordinate cycles are independent.  Therefore
\begin{equation}\label{eq:fixed}
F_r^H(g)=\prod_{C\in\operatorname{Cyc}(\pi^r)}
|\operatorname{Fix}_Q(h_{C,r})|.
\end{equation}
The exact vertex-cycle counts are then
\begin{equation}\label{eq:mobius}
c_\ell^H(g)=\frac1\ell\sum_{d\mid\ell}\mu(\ell/d)F_d^H(g).
\end{equation}
Substitution in \eqref{eq:master} proves the proposed Hamming formula.  For
pure coordinate permutations the label products are identities, reducing
\eqref{eq:fixed} to $q^{c(\pi^r)}$.

Two ordered pairs of words are equivalent exactly when they have equal
Hamming distance.  Hence vertex transitivity and the distances $0,\ldots,D$
give
\[
[m_{(1)}]=1,\qquad [m_{(2)}]=[m_{(1,1)}]=D+1.
\]

\section{Independent computational checks}
The implementation explicitly constructs every word permutation and its
$q^D\times q^D$ matrix $P$.  It computes complete symmetric characters from
actual matrix traces,
\[
d h_d(P)=\sum_{r=1}^d\operatorname{tr}(P^r)h_{d-r}(P),\quad h_0=1,
\]
and averages $\prod_jh_{\tau_j}(P)$.  A second route extracts coefficients
from \eqref{eq:master}; a third explicitly traverses the orbits in
\eqref{eq:orbit}.  Separately, the coordinate maps and alphabet labels are
powered by composition, \eqref{eq:fixed} is evaluated, and \eqref{eq:mobius}
is required to reproduce the actual vertex cycles for every group element.

\section{Exact results}
Values are ordered by
$\tau=(1),(2),(1,1),(3),(2,1),(1,1,1)$.
\begin{center}
\begin{tabular}{@{}lrrl@{}}
\toprule
Action & $|G|$ & $|X|$ & common coefficient vector\\
\midrule
$S_2\wr S_2$ on $Q^2$ & 8 & 4 & $(1,3,3,4,7,10)$\\
$S_2\wr S_3$ on $Q^3$ & 48 & 8 & $(1,4,4,7,13,20)$\\
$S_3\wr S_2$ on $Q^2$ & 72 & 9 & $(1,3,3,7,11,15)$\\
\bottomrule
\end{tabular}
\end{center}
All 18 matrix/cycle/orbit comparisons and every fixed-word reconstruction
pass.  For $(q,D)=(3,2)$ at $(t_1,t_2)=(0.037,0.061)$, the determinant and
cycle averages are $1.1254339873575003$ and $1.1254339873575000$; the absolute
error is $2.22\times10^{-16}$.

\section{Reproduction}
Run
\begin{verbatim}
marimo run marimo_notebooks/hamming_scheme.py
\end{verbatim}
from the repository root.  The table in the notebook exposes every exact
comparison and the global fixed-point reconstruction result.
\end{document}
