\documentclass[11pt]{article}
\usepackage[margin=0.84in]{geometry}
\usepackage{amsmath,amssymb,mathtools,booktabs,microtype,xcolor,fancyhdr,hyperref}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}
\lhead{$S_n$ on $k$-subsets}\rhead{Proof and exhaustive verification}\cfoot{\thepage}
\newcommand{\Fix}{\operatorname{Fix}}
\newcommand{\SMG}{\operatorname{SMG}}
\title{\color{navy}\bfseries The Non-Poisson $\Lambda$-Limit for $k$-Subsets\\
\large Exact matrix formula, proof, and exhaustive small-rank checks}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle

\begin{abstract}
For fixed $k$, we prove the joint limit of all power traces of the permutation
representation of $S_n$ on $k$-subsets.  The limit is an explicit functional
of independent $Z_j\sim\mathrm{Poisson}(1/j)$ variables and is non-Poisson for
$k\ge2$.  An independent program constructs 1,584 representation matrices in
dimensions $6,10,15,20$ and performs 6,336 exact power-trace comparisons.
\end{abstract}

\section{The representation and exact formula}
Let $\Omega_{n,k}=\binom{[n]}k$ and let $\rho_{n,k}$ permute its standard
basis.  Since the trace of a permutation matrix counts fixed basis vectors,
\[
 p_r(X_{\rho}(\sigma))=\operatorname{Tr}(\rho(\sigma)^r)
 =|\Fix_{\Omega_{n,k}}(\sigma^r)|.
\]
Write $A_j(\sigma)$ for the number of $j$-cycles of $\sigma$.  A $j$-cycle
splits under the $r$th power into $d=\gcd(j,r)$ cycles of length $j/d$.
An invariant subset independently includes or excludes each resulting cycle,
so
\begin{equation}\label{eq:fixed}
\boxed{F_{k,r}(\sigma)=[u^k]\prod_{j=1}^{kr}
\left(1+u^{j/\gcd(j,r)}\right)^{\gcd(j,r)A_j(\sigma)}.}
\end{equation}
The cutoff is valid: if $j>kr$, then $j/\gcd(j,r)>k$.

For a partition $\lambda$, direct averaging therefore gives
\begin{equation}\label{eq:smg}
\SMG_{n,k}=\sum_\lambda\frac{\mathbb E_{S_n}
[\prod_aF_{k,\lambda_a}(\sigma)]}{z_\lambda}p_\lambda.
\end{equation}

\section{Coefficientwise limit}
Fix a finite set of exponents.  Every variable in \eqref{eq:fixed} is a
polynomial in finitely many cycle counts.  We use the following elementary
factorial-moment lemma.  For fixed nonnegative integers $r_1,\ldots,r_R$,
\[
\mathbb E\prod_{j=1}^R(A_j)_{r_j}=\prod_{j=1}^Rj^{-r_j}
\quad\text{whenever}\quad n\geq\sum_{j=1}^Rjr_j.
\]
Indeed, the falling-factorial product counts ordered selections of pairwise
disjoint distinguished cycles, with $r_j$ cycles of length $j$.  Choosing
their supports and cyclic orders and then permuting the remaining labels gives
$n!\prod_jj^{-r_j}/n!$.  These are exactly the joint factorial moments of
independent $Z_j\sim\operatorname{Poisson}(1/j)$.  Hence the finite vector of
cycle counts converges jointly, and every factorial moment required by a fixed
coefficient is eventually equal to its limiting value.  Thus
\[
(A_1,\ldots,A_R)\Longrightarrow(Z_1,\ldots,Z_R),\qquad
Z_j\ \text{independent},\quad Z_j\sim\mathrm{Poisson}(1/j).
\]
Thus
\[
F_{k,r}(\sigma)\Longrightarrow F^{(\infty)}_{k,r}:=
[u^k]\prod_{j=1}^{kr}
(1+u^{j/\gcd(j,r)})^{\gcd(j,r)Z_j},
\]
jointly for every finite collection of $r$'s.  Moment stabilization permits
expectations in \eqref{eq:smg} to pass to the limit, proving the coefficientwise
$\Lambda$-limit
\[
\boxed{\SMG^{(k)}_\infty=\sum_\lambda
\frac{\mathbb E[\prod_aF^{(\infty)}_{k,\lambda_a}]}{z_\lambda}p_\lambda.}
\]
For $r=1$, $F^{(\infty)}_{k,1}$ has mean one.  The action has rank $k+1$
once $n\ge2k$, hence its limiting second moment is $k+1$ and its variance is
$k$.  A Poisson variable of mean one has variance one; consequently the limit
is non-Poisson for $k\ge2$.  For example
\[
F^{(\infty)}_{2,1}=\binom{Z_1}{2}+Z_2.
\]

\section{Exact finite formula checks}
For each $(n,k)=(4,2),(5,2),(6,2),(6,3)$, the verifier enumerates all $n!$
permutations, induces their action on lexicographically ordered $k$-subsets,
and materializes the corresponding zero-one matrix.  Distinct byte keys prove
that the constructed matrices form a faithful copy of $S_n$.  It then compares
the integer matrix traces $\operatorname{Tr}(M^r)$, $1\le r\le4$, with
\eqref{eq:fixed}.  Six joint products indexed by
$(1),(2),(3),(1,1),(2,1),(1,1,1)$ are averaged independently on both sides.

\begin{center}
\begin{tabular}{@{}rrrrrr@{}}\toprule
$n$&$k$&dimension&matrices&trace checks&$(\mathbb E F,\operatorname{Var}F)$\\\midrule
4&2&6&24&96&$(1,2)$\\
5&2&10&120&480&$(1,2)$\\
6&2&15&720&2880&$(1,2)$\\
6&3&20&720&2880&$(1,3)$\\\bottomrule
\end{tabular}
\end{center}
All 6,336 trace comparisons and all 24 joint-average comparisons pass exactly.
These exhaustive matrices verify the finite fixed-point identity, not the
asymptotic convergence; convergence is proved by the factorial-moment lemma
above.

\section{Reproduction and reference}
\begin{verbatim}
.venv/bin/python lambda_distributions/proofs/limiting_lambda_claims/
  symmetric_k_subsets/verification.py
.venv/bin/marimo edit marimo_notebooks/symmetric_k_subsets.py
\end{verbatim}
Join each wrapped command onto one line.  For the fixed-point and moment
background, see P.~Diaconis, J.~Fulman, and R.~Guralnick,
\emph{On fixed points of permutations},
\url{https://arxiv.org/abs/0708.2643}.
\end{document}
