\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,booktabs,array,microtype}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue,urlcolor=blue]{hyperref}
\newcommand{\Sym}{\operatorname{Sym}}
\newcommand{\Fix}{\operatorname{Fix}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\cS}{\mathcal S}
\title{Johnson Schemes: Permutation Molien Formula\\Proof and Exact Verification}
\author{Executable companion note}
\date{17 July 2026}
\begin{document}
\maketitle
\begin{abstract}
For the permutation representation on the $k$-subsets of $[v]$, we prove the
cycle-factor and orbit formulas, derive the fixed-subset formula, treat the
middle-dimensional complementation coset, and verify all coefficients through
total degree three in three representative finite cases.  The computation uses
the actual permutation matrices and does not assume the desired cycle formula.
\end{abstract}

\section{Master identity}
Let a finite group $G$ act on a finite set $X$, let $V=\mathbb C[X]$, and let
$c_\ell(g)$ count the $\ell$-cycles of $g$ on $X$.  Then
\begin{equation}\label{eq:master}
 \cS_{G,X}(\mathbf t)=\frac1{|G|}\sum_{g\in G}
 \prod_{i\ge1}\prod_{\ell\ge1}(1-t_i^\ell)^{-c_\ell(g)}.
\end{equation}
Indeed, the eigenvalues of an $\ell$-cycle are all the $\ell$th roots of
unity, so $\det(I-tP_\ell)=1-t^\ell$.  Multiplication over disjoint cycles and
the finite-group Molien average prove \eqref{eq:master}.

For a partition $\tau=(\tau_1,\ldots,\tau_s)$, $\Sym^{\tau_j}(V)$ is the
permutation module on multisets of size $\tau_j$ in $X$.  Burnside's lemma
therefore gives the independent interpretation
\begin{equation}\label{eq:orbit}
 [m_\tau]\cS_{G,X}=\#\bigl(G\backslash
 (\Sym^{\tau_1}X\times\cdots\times\Sym^{\tau_s}X)\bigr).
\end{equation}

\section{Johnson fixed points and complementation}
Put $X=\binom{[v]}k$.  If $\sigma\in S_v$ has $a_j$ cycles of length $j$,
then a $j$-cycle splits under $\sigma^r$ into $\gcd(j,r)$ cycles of length
$j/\gcd(j,r)$.  A fixed subset is exactly a union of these cycles; hence
\begin{equation}\label{eq:johnsonfixed}
 F_r^J(\sigma)=[u^k]\prod_{j\ge1}
 \left(1+u^{j/\gcd(j,r)}\right)^{a_j\gcd(j,r)}.
\end{equation}
The vertex-cycle counts follow from
\begin{equation}\label{eq:mobius}
 c_\ell(\sigma)=\frac1\ell\sum_{d\mid\ell}
 \mu(\ell/d)F_d^J(\sigma).
\end{equation}

When $v=2k$ and $k>1$, complementation $T(A)=A^c$ commutes with $S_{2k}$
and supplies the additional coset in the full Johnson-scheme automorphism
group.  For odd $r$, a subset fixed by $\sigma^rT$ must alternate membership
around every cycle of $\sigma^r$.  Thus
\begin{equation}\label{eq:coset}
F_r(\sigma T)=\begin{cases}
F_r^J(\sigma),&r\ \text{even},\\
2^{c(\sigma^r)},&r\ \text{odd and every cycle of $\sigma^r$ is even},\\
0,&\text{otherwise}.
\end{cases}
\end{equation}
Every alternating choice has size exactly half of $v$, so no further size
constraint is missing.  Equations \eqref{eq:master}--\eqref{eq:coset} prove
the claimed formula, including the middle case.

\section{Verification design}
For each group element, the code explicitly builds its $|X|\times|X|$
permutation matrix $P$.  It computes $h_d(P)=\operatorname{tr}(\Sym^dP)$ from
the matrix traces by Newton's recurrence
\[
 d h_d(P)=\sum_{r=1}^d\operatorname{tr}(P^r)h_{d-r}(P),\qquad h_0=1.
\]
It averages $\prod_j h_{\tau_j}(P)$ over the group.  Independently, it
extracts the same coefficient from the cycle factors in \eqref{eq:master}
and counts the orbits in \eqref{eq:orbit} by explicit set traversal.  Finally,
it evaluates \eqref{eq:johnsonfixed} or \eqref{eq:coset} for every required
power and checks that \eqref{eq:mobius} reproduces every actual cycle count.

\section{Exact results}
The entries list the common value in the order
$\tau=(1),(2),(1,1),(3),(2,1),(1,1,1)$.
\begin{center}
\begin{tabular}{@{}lrrl@{}}
\toprule
Action & $|G|$ & $|X|$ & common coefficient vector\\
\midrule
$\Aut J(4,2)$ (including $T$) & 48 & 6 & $(1,3,3,5,8,11)$\\
$S_5$ on $2$-subsets & 120 & 10 & $(1,3,3,7,11,15)$\\
$S_5$ on $1$-subsets & 120 & 5 & $(1,2,2,3,4,5)$\\
\bottomrule
\end{tabular}
\end{center}
All 18 triple comparisons are exact, and all fixed-point reconstructions pass.

At $(t_1,t_2)=(0.037,0.061)$ the direct determinant average and cycle product
for $\Aut J(4,2)$ coincide at
\[
1.1237420509116285\qquad\text{(reported error $0$)}.
\]
The pair values also give $[m_{(2)}]=[m_{(1,1)}]=\min(k,v-k)+1$, as predicted.

\section{Reproduction}
From the repository root run
\begin{verbatim}
marimo run marimo_notebooks/johnson_scheme.py
\end{verbatim}
or run the family module directly with Python.  The notebook displays every
matrix/cycle/orbit comparison rather than only the summary above.
\end{document}
