\documentclass[11pt]{article}
\usepackage[margin=0.84in]{geometry}
\usepackage{amsmath,amssymb,mathtools,booktabs,microtype,xcolor,fancyhdr,hyperref}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}
\lhead{Barnes--Wall family II}\rhead{Real Clifford normalizer}\cfoot{\thepage}
\newcommand{\MGF}{\mathfrak M}
\title{\color{navy}\bfseries The Real Clifford Layer of Barnes--Wall Symmetry\\
\large Coset reduction, code invariants, and explicit normalizer checks}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle

\begin{abstract}
We prove the exact reduction of the real Clifford sigma--MGF to coset
averages over $O^+(2m,2)$ and state precisely what the self-dual-code theorem
does and does not imply for mixed coefficients.  Exhaustive matrix closures
in dimensions $2$ and $4$ recover group orders $16$ and $2304$, quotient
orders $2$ and $72$, lift independence, and 16 direct coefficient matches.
\end{abstract}

\section{Exact quotient formula}
Let $E_m\triangleleft\mathcal C_m=N_{O(V_m)}(E_m)$, with
\[
 \mathcal C_m/E_m\cong O^+(2m,2).
\]
For $A$ in the quotient and any lift $\widetilde A$, define
\[
 \Phi_m(A;\mathbf t)=\frac1{|E_m|}\sum_{e\in E_m}
 \prod_j\det(I-t_je\widetilde A)^{-1}.
\]
Replacing $\widetilde A$ by $e_0\widetilde A$ permutes $E_m$, so $\Phi_m$
is well defined.  Partitioning $\mathcal C_m$ into right cosets proves
\begin{equation}\label{eq:coset}
\boxed{\MGF_{\mathcal C_m,m}(\mathbf t)=
\frac1{|O^+(2m,2)|}\sum_{A\in O^+(2m,2)}\Phi_m(A;\mathbf t).}
\end{equation}
Coefficient extraction gives
\[
[m_\tau]\MGF_{\mathcal C_m,m}=
\dim\left(\bigotimes_i\operatorname{Sym}^{\tau_i}V_m\right)^{\mathcal C_m}.
\]

\section{One-row code theorem and its boundary}
For a binary self-dual code $C\leq\mathbb F_2^{2k}$, its genus-$m$
complete weight enumerator is
\[
 \operatorname{cwe}_m(C)=\sum_{(c^{(1)},\ldots,c^{(m)})\in C^m}
 \prod_{j=1}^{2k}x_{(c_j^{(1)},\ldots,c_j^{(m)})}.
\]
The Clifford invariant theorem identifies the degree-$2k$ invariant space
with the span of these enumerators.  In the stable range $m\geq k-1$, they
are independent and therefore
\begin{equation}\label{eq:codes}
 [m_{(2k)}]\MGF_{\mathcal C_m,m}
 =\#\{\text{binary self-dual codes of length }2k\}/\!\cong.
\end{equation}
Consequently the stable values in degrees $2,4,6,8$ are $1,1,1,2$.
The identification with genus-$m$ complete weight enumerators and their
independence in this range are external inputs from the
Nebe--Rains--Sloane invariant theorem cited below.  This note proves the coset
formula and verifies low-rank examples; it does not reprove that theorem.
The degree-eight value is guaranteed by \eqref{eq:codes} only for $m\geq3$;
our smaller closures also happen to give two, but we report that as a direct
low-rank computation rather than misuse the stable-range hypothesis.

For a general partition $\tau$, block-symmetrizing the full code tensor
produces explicit invariants.  The ordinary invariant-ring theorem does not
by itself show that these span every mixed fixed space.  Thus \eqref{eq:coset}
is exact for all $\tau$, while a code-indexed equality for all mixed $\tau$
would require a separate multigraded theorem.

\section{Explicit matrices and exhaustive checks}
We generate the real Pauli matrices together with tensor placements of
\[
 H=2^{-1/2}\begin{pmatrix}1&1\\1&-1\end{pmatrix}
\]
and, for $m=2$, CNOT.  Breadth-first closure yields
\[
|\mathcal C_1|=16=8\cdot2,
\qquad |\mathcal C_2|=2304=32\cdot72.
\]
Every closed matrix is checked to be real orthogonal.  Conjugation by every
selected coset representative sends the Pauli group to itself.  The program
then partitions the closed group into Pauli cosets, averages each coset, and
compares \eqref{eq:coset} with an independent direct average.  Each lift is
also multiplied by a varying Pauli matrix, and every corresponding pair of
coset contributions is compared before the quotient average is taken.

\begin{center}
\begin{tabular}{@{}cccrrr@{}}\toprule
$m$&$|\mathcal C_m|$&$|O^+|$&$\tau$&direct&coset\\\midrule
1&16&2&$(2)$&1&1\\
1&16&2&$(8)$&2&2\\
1&16&2&$(2,2)$&2&2\\
2&2304&72&$(4)$&1&1\\
2&2304&72&$(8)$&2&2\\
2&2304&72&$(4,2)$&2&2\\\bottomrule
\end{tabular}
\end{center}
The full sweep tests eight partitions at each rank.  Stable-code references
are applied only when $m\geq k-1$; all 16 rows pass.

\section{Reproduction and references}
\begin{verbatim}
.venv/bin/python lambda_distributions/proofs/barnes_wall_lattice_automorphism_groups/
  real_clifford/verification.py
.venv/bin/marimo edit marimo_notebooks/real_clifford.py
\end{verbatim}
Join wrapped commands onto one line.  For the invariant theorem, see
G.~Nebe, E.~Rains, and N.~J.~A.~Sloane,
\emph{The invariants of the Clifford groups},
\url{https://arxiv.org/abs/math/0001038}.
\end{document}
