\documentclass[11pt]{article}
\usepackage[margin=0.84in]{geometry}
\usepackage{amsmath,amssymb,mathtools,booktabs,microtype,xcolor,fancyhdr,hyperref}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}
\lhead{Barnes--Wall family III}\rhead{Lattice automorphism subgroup}\cfoot{\thepage}
\newcommand{\MGF}{\mathfrak M}
\title{\color{navy}\bfseries Barnes--Wall Lattice Automorphism Groups\\
\large The index-two correction and direct lattice-preservation verification}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle

\begin{abstract}
Away from the exceptional rank $m=3$, the Barnes--Wall automorphism group is
an index-two subgroup of the real Clifford group.  We prove the resulting
twisted sigma--MGF identity and verify it exhaustively in the concrete models
$BW_2=\mathbb Z^2$ and $BW_4=D_4$.  The constructed groups have orders $8$
and $1152$ and agree exactly with the Clifford matrices preserving the
specified lattice bases.  All 16 coefficient comparisons pass.
\end{abstract}

\section{The index-two identity}
Assume $m\ne3$ and write
\[
B_m=\operatorname{Aut}(BW_{2^m})=\ker\varepsilon_m
\quad\text{inside}\quad \mathcal C_m,
\]
where $\varepsilon_m:\mathcal C_m\to\{\pm1\}$ is the nontrivial quotient
character.  Define the twisted series
\[
\MGF^{\varepsilon_m}_{\mathcal C_m,m}(\mathbf t)
=\frac1{|\mathcal C_m|}\sum_{g\in\mathcal C_m}\varepsilon_m(g)
\prod_j\det(I-t_jg)^{-1}.
\]
Since $\mathbf1_{B_m}(g)=(1+\varepsilon_m(g))/2$ and
$|B_m|=|\mathcal C_m|/2$, direct substitution proves
\begin{equation}\label{eq:twist}
\boxed{\MGF_{B_m,m}=\MGF_{\mathcal C_m,m}
+\MGF^{\varepsilon_m}_{\mathcal C_m,m}.}
\end{equation}
For $W_\tau=\bigotimes_i\operatorname{Sym}^{\tau_i}V_m$, coefficient
extraction equivalently gives
\[
\dim W_\tau^{B_m}=\dim W_\tau^{\mathcal C_m}
+\dim\operatorname{Hom}_{\mathcal C_m}(\varepsilon_m,W_\tau).
\]
The second nonnegative summand is precisely the lattice-sensitive sector.

The structural equality
$\operatorname{Aut}(BW_{2^m})=\ker\varepsilon_m\leq\mathcal C_m$ and the
exceptional behavior at $m=3$ are external inputs from the cited
Barnes--Wall/Clifford-group literature.  Equation \eqref{eq:twist} is proved
conditionally from this index-two statement, and the verifier checks the
concrete ranks $m=1,2$.

\section{Concrete lattice models}
For $m=1$ we use $BW_2=\mathbb Z^2$.  Its automorphism group is the signed
permutation group of order $8$, equal to $E_1$, while the real Clifford group
has order $16$.

For $m=2$ we use
\[
 BW_4=D_4=\{x\in\mathbb Z^4:\textstyle\sum_i x_i\equiv0\pmod2\}.
\]
A basis is given by the columns of
\[
L=\begin{pmatrix}
1&0&0&0\\-1&1&0&0\\0&-1&1&1\\0&0&-1&1
\end{pmatrix}.
\]
Starting with $E_2$, tensor Hadamard $H\otimes H$, CNOT, and controlled-$Z$,
matrix closure gives a subgroup of order $1152$.  Independently, for each of
the $2304$ matrices $g\in\mathcal C_2$ the verifier tests whether
$L^{-1}gL\in GL(4,\mathbb Z)$.  Exactly the same $1152$ matrices pass.  Thus
the subgroup identification uses lattice preservation, not only an expected
order.  Every matrix in both closed groups is also checked to have real
entries and satisfy $g^{\mathsf T}g=I$.

\section{Direct versus twisted averages}
For each partition $\tau$, the direct side averages
$\prod_i h_{\tau_i}(\operatorname{spec}g)$ over $B_m$.  On the right side,
the same integrand is averaged first with weight $1$ and then with weight
$\varepsilon_m(g)$ over all of $\mathcal C_m$.  Membership in the independently
identified lattice stabilizer determines the sign.

\begin{center}
\begin{tabular}{@{}cccrrrrr@{}}\toprule
$m$&$|B_m|$&$\tau$&direct&Clifford&twisted&sum\\\midrule
1&8&$(4)$&2&1&1&2\\
1&8&$(8)$&3&2&1&3\\
1&8&$(4,2)$&4&2&2&4\\
2&1152&$(4)$&1&1&0&1\\
2&1152&$(6)$&2&1&1&2\\
2&1152&$(8)$&3&2&1&3\\
2&1152&$(4,2)$&3&2&1&3\\\bottomrule
\end{tabular}
\end{center}
These rows exhibit a real distinction: the first tested one-row twisted
contribution occurs in degree $4$ for $B_1$ but only degree $6$ for $B_2$.
The full sweep tests eight partitions at each rank, including mixed powers;
all 16 instances of \eqref{eq:twist} pass.

\section{Scope and reproduction}
No $m=3$ conclusion is drawn: that rank is explicitly exceptional to the
index-two statement.  Ranks $m\geq4$ are covered algebraically by
\eqref{eq:twist}, but exhaustive closure is not a reasonable verification
strategy because the groups grow rapidly.
\begin{verbatim}
.venv/bin/python lambda_distributions/proofs/barnes_wall_lattice_automorphism_groups/
  barnes_wall/verification.py
.venv/bin/marimo edit marimo_notebooks/barnes_wall.py
\end{verbatim}
Join each wrapped command onto one line.  Background on the group chain and
the exceptional rank appears in G.~Nebe, E.~Rains, and N.~J.~A.~Sloane,
\emph{The invariants of the Clifford groups},
\url{https://arxiv.org/abs/math/0001038}.
\end{document}
