\documentclass[11pt]{article}
\usepackage[margin=0.85in]{geometry}
\usepackage{amsmath,amssymb,mathtools,booktabs,microtype,xcolor,fancyhdr,hyperref}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}\definecolor{red}{HTML}{9B2C2C}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}\lhead{Scalar-extended Pauli groups}\rhead{Proof and verification}\cfoot{\thepage}
\newcommand{\Sgm}{\mathcal S}
\title{\color{navy}\bfseries Scalar-Extended Pauli Sigma--MGFs\\\large Proof, qubit recovery, and sign correction}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle
\begin{abstract}
We prove the coefficient form of the scalar-extended Pauli formula and test
the complete matrix groups $\mu_4P_{2,1}$, $\mu_4P_{2,2}$, and
$\mu_6P_{3,1}$.  All 13 representative checks pass.  The calculation also
locates a sign error in the supplied qubit-recovery prose: equation (13)
specializes with a positive, projected noncentral sector, not a minus sign.
\end{abstract}

\section{Formula and proof}
Let $q=p^f$, $N=q^n$, and let $s$ be divisible by $p$.  For the finite lift
$\widetilde P_{q,n}^{(s)}=\mu_sP_{q,n}$, its distinct matrices may be indexed
by $\alpha\in\mu_s$ and $v\in\mathbb F_q^{2n}$, so its order is $sq^{2n}$.
For odd $p$ use the usual phase-normalized Weyl operators.  For qubits we use
$\{I,X,Z,iXZ\}$ on each tensor factor and adjoin $\mu_4$.

If $v=0$, the matrix is $\alpha I_N$.  If $v\ne0$, the $p$ spectral branches
of the phase-normalized $W(v)$ occur with multiplicity $N/p$, and therefore
\[
\det(I-t\alpha W(v))=(1-\alpha^p t^p)^{N/p}.
\]
Averaging these two sectors proves
\begin{equation}\label{eq:scalar}
\Sgm_{\widetilde P}(\mathbf t)=
q^{-2n}\frac1s\sum_{\alpha\in\mu_s}\prod_j(1-\alpha t_j)^{-N}
+(1-q^{-2n})\frac1s\sum_{\alpha\in\mu_s}
\prod_j(1-\alpha^pt_j^p)^{-N/p}.
\end{equation}
Root-of-unity orthogonality forces $s\mid|\tau|$ in both sectors.  Thus,
with
\[
A_\tau=\prod_i\binom{N+\tau_i-1}{\tau_i},\qquad
B_\tau=\mathbf1_{p\mid\tau_i\ \forall i}
\prod_i\binom{N/p+\tau_i/p-1}{\tau_i/p},
\]
the coefficient formula is
\begin{equation}\label{eq:coeff}
[m_\tau]\Sgm_{\widetilde P}=\begin{cases}
0,&s\nmid|\tau|,\\
q^{-2n}A_\tau+(1-q^{-2n})B_\tau,&s\mid|\tau|.
\end{cases}
\end{equation}

\section{Qubit specialization and correction}
For $p=q=2$, $s=4$, equation \eqref{eq:scalar} gives
\[
4^{-n}\Pi_4\prod_j(1-t_j)^{-2^n}
+(1-4^{-n})\frac14\sum_{\alpha\in\mu_4}
\prod_j(1-\alpha^2t_j^2)^{-2^{n-1}}.
\]
The second summand is an average of determinant reciprocals and has positive
coefficients in allowed multidegrees.  In particular, for one qubit,
\[
[t^4]\Sgm=\tfrac14\binom{5}{4}+\tfrac34\binom{2}{2}=2.
\]
The supplied “qubit recovery” paragraph instead writes a minus before the
noncentral term, which would give $1/2$, not even an integer invariant
dimension.  Therefore that minus sign cannot be correct.  Equation (13)
itself, interpreted as the root-of-unity average above, is consistent with
the direct matrices and with \eqref{eq:coeff}.

\section{Verification strategy and results}
For each matrix the verifier computes the coefficients $h_d$ of
$\det(I-tg)^{-1}$ from its eigenvalues, then independently averages
$\prod_i h_{\tau_i}(g)$ over every group matrix.  It compares this character
average with \eqref{eq:coeff}.  Representative results are:
\begin{center}
\begin{tabular}{@{}lrrr@{}}\toprule
group & $\tau$ & direct & formula\\\midrule
$\mu_4P_{2,1}$ & $(2)$&0&0\\
$\mu_4P_{2,1}$ & $(4)$&2&2\\
$\mu_4P_{2,1}$ & $(2,2)$&3&3\\
$\mu_4P_{2,2}$ & $(4)$&5&5\\
$\mu_4P_{2,2}$ & $(4,4)$&85&85\\
$\mu_6P_{3,1}$ & $(3)$&0&0\\
$\mu_6P_{3,1}$ & $(6)$&4&4\\
$\mu_6P_{3,1}$ & $(5,1)$&7&7\\\bottomrule
\end{tabular}
\end{center}
All 13 built-in comparisons pass to $2\times10^{-8}$.  The cases $(2)$ for
$s=4$ and $(3)$ for $s=6$ directly exercise the stronger scalar selection
rule, while $(3,1)$ and $(5,1)$ test mixed symmetric powers.

\section{Reproduction}
\begin{verbatim}
.venv/bin/python for_this_guy/quantum_matrix_groups/
  scalar_extended_pauli/verification.py
.venv/bin/marimo edit marimo_notebooks/scalar_extended_pauli.py
\end{verbatim}
Join each displayed command onto one shell line.
\end{document}
