\documentclass[11pt]{article}
\usepackage[margin=0.82in]{geometry}
\usepackage{amsmath,amssymb,booktabs,microtype,xcolor,fancyhdr,hyperref}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}
\lhead{$Co_0$ Leech representation}\rhead{Proof and verification}\cfoot{\thepage}
\newcommand{\Sgm}{\mathcal S}
\title{\color{navy}\bfseries The $Co_0$ Leech-Space Sigma--MGF\\
\large Frame-shape proof, central parity, and a 24-dimensional stress test}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle

\begin{abstract}
We prove the frame-shape class formula for the natural $24$-dimensional
complexified Leech-lattice representation of $Co_0$ and the vanishing of all
odd total-degree coefficients.  Since direct enumeration of $Co_0$ is not a
reasonable notebook computation, we test the determinant mechanism at the
correct dimension on the explicit signed-coordinate subgroup
$\{\pm P_g:g\in M_{24}\}$.  Direct invariant counts and exact class expansion
agree through degree six.  The distinction between this stress test and a
full $Co_0$ class-table evaluation is kept explicit.
\end{abstract}

\section{Frame-shape formula}
Let $V=\Lambda_{\rm Leech}\otimes_{\mathbb Z}\mathbb C$.  If a conjugacy-class
representative $g\in Co_0$ has frame shape
$\pi_g=\prod_{r\geq1}r^{k_r(g)}$, by definition
\[
\det(xI-g)=\prod_{r\geq1}(x^r-1)^{k_r(g)},\qquad
\sum_r r k_r(g)=24.
\]
Substitute $x=t^{-1}$ and multiply by $t^{24}$ to obtain
\[
\det(I-tg)=\prod_{r\geq1}(1-t^r)^{k_r(g)}.
\]
The generalized Molien formula, grouped by conjugacy class, is therefore
\begin{equation}\label{eq:co0}
\boxed{\Sgm_{Co_0,V}(\mathbf t)=
\sum_{C\subseteq Co_0}\frac{|C|}{|Co_0|}
\prod_{i,r}(1-t_i^r)^{-k_r(C)}.}
\end{equation}
Negative frame exponents cause no difficulty: each class contribution is a
rational function whose product is the characteristic determinant.

\section{Central parity theorem}
The central element $-I$ belongs to $Co_0$.  On
$\operatorname{Sym}^{\tau}V=\bigotimes_j\operatorname{Sym}^{\tau_j}V$ it
acts by $(-1)^{|\tau|}$.  A fixed vector in odd total degree would satisfy
$v=-v$, hence vanish.  Thus
\begin{equation}\label{eq:parity}
\boxed{[m_\tau]\Sgm_{Co_0,V}=0\quad\text{whenever }|\tau|\text{ is odd}.}
\end{equation}
In particular $[m_{(1)}]=0$, unlike the $M_{24}$ permutation representation,
whose all-ones vector is fixed.  The two 24-dimensional series therefore
differ in their first positive degree.

This proof is representation theoretic and does not require any frame-shape
table.  As a data-integrity check on a full table, the class contributions in
every odd degree must cancel exactly.

\section{A tractable subgroup at the true dimension}
In the standard coordinate construction of the Leech lattice, its monomial
automorphism group contains $M_{24}$ coordinate permutations and the central
sign.  We use
\[
H=\{\pm P_g:g\in M_{24}\}\leq Co_0.
\]
The notebook explicitly constructs three $24\times24$ integer orthogonal
generators $P_a,P_b,-I$, using the ATLAS permutations $a,b$ documented in the
companion $M_{24}$ note.  This is large enough to exercise the actual ambient
dimension while retaining an exact aggregated class computation.

If $g$ has cycle counts $c_r(g)$, then
\[
\det(I-tP_g)=\prod_r(1-t^r)^{c_r(g)},\qquad
\det(I+tP_g)=\prod_r(1-(-t)^r)^{c_r(g)}.
\]
Averaging the two signs projects onto even total degree.  Hence
\begin{equation}\label{eq:signed}
\Sgm_H(t)=\frac1{2|M_{24}|}\sum_{g\in M_{24}}
\left\{\prod_r(1-t^r)^{-c_r(g)}+
\prod_r(1-(-t)^r)^{-c_r(g)}\right\}.
\end{equation}

\section{Independent verification}
For even $k$, $-I$ acts trivially on $\operatorname{Sym}^kV$, so the $H$-fixed
space has a basis indexed by $M_{24}$ orbits of degree-$k$ multisets.  For odd
$k$, \eqref{eq:parity} makes the dimension zero.  This gives a direct
generator-orbit algorithm independent of the class expansion
\eqref{eq:signed}.  Exact results are:
\begin{center}
\begin{tabular}{@{}crrc@{}}\toprule
$k$ & direct invariant dimension & signed class formula & parity\\\midrule
1&0&0&odd\\
2&2&2&even\\
3&0&0&odd\\
4&5&5&even\\
5&0&0&odd\\
6&12&12&even\\\bottomrule
\end{tabular}
\end{center}
The degree-six value also exercises the first point at which the underlying
$M_{24}$ multiset orbits differ from those of $S_{24}$.

\section{What this does and does not verify}
The proof of \eqref{eq:co0} is exact for full $Co_0$, and the proof of
\eqref{eq:parity} establishes every odd-degree zero for full $Co_0$.  The
matrix sweep validates determinant expansion, sign cancellation, and direct
invariant counting in dimension $24$ on a genuine subgroup.  It does not
claim that the even coefficients $2,5,12$ are full-$Co_0$ coefficients; fixed
spaces can shrink when the acting group grows.  A full numerical evaluation
of \eqref{eq:co0} requires the complete $Co_0$ frame-shape and class-size
table, which is deliberately not reconstructed from partial data.

Run
\begin{verbatim}
.venv/bin/python for_this_guy/sporadic_matrix_groups/
  co0_leech/verification.py
.venv/bin/marimo edit marimo_notebooks/co0_leech.py
\end{verbatim}
after joining wrapped lines.  For background on the Leech representation and
Conway fixed lattices, see \url{https://arxiv.org/abs/1505.06420}.
\end{document}
