\documentclass[11pt]{article}
\usepackage[margin=0.82in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs,tabularx}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true,linkcolor=NavyBlue,urlcolor=BrickRed]{hyperref}
\usepackage{microtype}
\usepackage{fancyhdr}
\usepackage{enumitem}

\definecolor{ink}{HTML}{17302A}
\definecolor{accent}{HTML}{287A62}
\definecolor{pale}{HTML}{EDF7F3}
\definecolor{passgreen}{HTML}{16794A}
\color{ink}
\pagestyle{fancy}
\fancyhf{}
\lhead{Representation variants}
\rhead{Proof and verification}
\cfoot{\thepage}
\setlength{\headheight}{14pt}
\newcommand{\Sfun}{\mathcal S}
\newcommand{\Sym}{\operatorname{Sym}}
\newcommand{\Reg}{\operatorname{Reg}}
\newcommand{\PASS}{\textcolor{passgreen}{\textbf{PASS}}}
\newcommand{\summarybox}[1]{\begin{center}\fcolorbox{accent}{pale}{\parbox{0.91\linewidth}{#1}}\end{center}}

\title{\textbf{Changing the Representation}\\[3pt]
\large Exterior powers and regular representations in the master formula}
\author{Computational verification note}
\date{17 July 2026}

\begin{document}
\maketitle

\summarybox{\textbf{Outcome.}  Two structurally different substitutions into
the finite-group master formula were tested.  For $\Lambda^2$ of the standard
three-dimensional $S_4$ representation, explicit compound matrices agree with
the pairwise-eigenvalue formula.  For the regular representation of $S_3$,
explicit $6\times6$ permutation matrices agree with the element-order formula.
All 38 multivariate coefficients through total degree five pass, as do direct
Reynolds projector checks.}

\section{The common master formula}

For any finite-dimensional representation $W$ of a finite group $G$,
\begin{equation}
\Sfun_{G,W}(\mathbf t)=\frac1{|G|}\sum_{g\in G}
\prod_{a\geq1}\det(I-t_a g|_W)^{-1}.                          \tag{1}
\end{equation}
The coefficient at $t_1^{\tau_1}\cdots t_\ell^{\tau_\ell}$ equals
\begin{equation}
\dim\left(\bigotimes_{a=1}^{\ell}\Sym^{\tau_a}W\right)^G,   \tag{2}
\end{equation}
because the coefficient is the group average of the tensor-product character,
which is the trace of the Reynolds projector.  Once (1) is established, a new
representation requires only a correct description of its eigenvalues.  This
report tests two such descriptions that use very different group data.

\section{Exterior powers}

If $g|_V$ has eigenvalues $\lambda_1,\ldots,\lambda_n$, then the wedge vectors
$v_{i_1}\wedge\cdots\wedge v_{i_k}$ show that $g|_{\Lambda^kV}$ has eigenvalues
\[
\lambda_I=\prod_{i\in I}\lambda_i,\qquad |I|=k.
\]
Therefore
\begin{equation}
\Sfun_{G,\Lambda^kV}(\mathbf t)=\frac1{|G|}\sum_{g\in G}
\prod_{a\geq1}\prod_{|I|=k}(1-t_a\lambda_I(g))^{-1}.         \tag{3}
\end{equation}
This proof needs no assumption that $G$ is a reflection group.

\subsection{Concrete test: $\Lambda^2$ of the $S_4$ reflection representation}

Start from the permutation representation of $S_4$ on $\mathbb C^4$ and
restrict it to
\[
V=\{(x_1,x_2,x_3,x_4):x_1+x_2+x_3+x_4=0\}.
\]
Numerical QR orthonormalization supplies a $4\times3$ basis matrix $B$, so the
explicit standard matrices are $B^*P_\pi B$.  The direct exterior-square
matrices are constructed without diagonalizing: in the ordered basis
$e_i\wedge e_j$, their entries are the $2\times2$ minors
\begin{equation}
(\Lambda^2g)_{(k,l),(i,j)}
=g_{ki}g_{lj}-g_{kj}g_{li}.                                  \tag{4}
\end{equation}
The formula route instead diagonalizes only the original $3\times3$ standard
matrices and forms the three pairwise eigenvalue products required by (3).
Thus an error in the compound-matrix construction is not reproduced in the
formula route.

\section{The regular representation}

Let $G$ act on its basis by left multiplication.  If $g$ has order $o(g)$,
then its permutation of the $|G|$ basis elements has $|G|/o(g)$ cycles, each of
length $o(g)$.  Hence
\begin{equation}
\det(I-tg|_{\Reg})=(1-t^{o(g)})^{|G|/o(g)}.                   \tag{5}
\end{equation}
If $N_m$ is the number of elements of order $m$, substitution into (1) gives
\begin{equation}
\Sfun_{G,\Reg}(\mathbf t)=\sum_m\frac{N_m}{|G|}
\prod_{a\geq1}(1-t_a^m)^{-|G|/m}.                            \tag{6}
\end{equation}

For coefficient extraction, use
\[
[t^d](1-t^m)^{-c}=
\begin{cases}
\binom{c+d/m-1}{d/m},&m\mid d,\\
0,&m\nmid d.
\end{cases}
\]
For $S_3$, the order counts are
\begin{equation}
N_1=1,\qquad N_2=3,\qquad N_3=2.                              \tag{7}
\end{equation}
Equations (6)--(7) therefore give an integer-arithmetic formula for every
multivariate coefficient.  The direct route constructs all six left-regular
$6\times6$ permutation matrices and never consults these order counts.

\section{Verification protocol}

For each representation and every partition $\tau$ through total degree five:
\begin{enumerate}[leftmargin=*,itemsep=3pt]
\item Construct all representation matrices explicitly.
\item Recursively compute $h_d$ from each matrix's eigenvalues and average
$\prod_a h_{\tau_a}$ over the group.
\item Compare with the representation-specific formula: pairwise spectra in
(3) for $S_4$, or binomial coefficient extraction from (6) for $S_3$.
\item Embed the normalized occupancy basis of $\Sym^2(W)$ in $W^{\otimes2}$,
form the explicit Reynolds matrix, and compare rank with the coefficient.
\end{enumerate}

\section{Results}

\begin{center}
\begin{tabularx}{\linewidth}{@{}Xrrrrc@{}}
\toprule
Representation & $|G|$ & $\dim W$ & degree & checks & result\\
\midrule
$S_4$ on $\Lambda^2(V_{\mathrm{std}})$ & 24 & 3 & $0$--$5$ & 19 & \PASS\\
$S_3$ on $\Reg$ & 6 & 6 & $0$--$5$ & 19 & \PASS\\
\bottomrule
\end{tabularx}
\end{center}

Selected matched dimensions are
\begin{center}
\begin{tabular}{@{}lrrrrrr@{}}
\toprule
Representation & $(1)$ & $(2)$ & $(3)$ & $(2,1)$ & $(2,2)$ & $(3,2)$\\
\midrule
$\Lambda^2(V_{\mathrm{std}})$ of $S_4$ & 0 & 1 & 0 & 0 & 3 & 1\\
$\Reg$ of $S_3$ & 1 & 5 & 10 & 21 & 78 & 196\\
\bottomrule
\end{tabular}
\end{center}
The large mixed coefficients in the regular case provide a useful stress test:
agreement is not caused merely by both routes rounding small numbers to zero.

For the exterior case, $\Sym^2(W)$ has dimension six and the Reynolds
projector has rank $1$, trace $0.9999999999999992$, and idempotence error
$8.31\times10^{-16}$.  For the regular case, $\Sym^2(W)$ has dimension $21$;
the projector has rank $5$, trace $4.999999999999999$, and idempotence error
$3.72\times10^{-16}$.  These ranks agree with the $(2)$ coefficient in each
row of the table.

\section{What these tests establish}

The exterior test exercises a nonlinear eigenvalue transformation and a
coordinate-level compound-matrix construction.  The regular test exercises a
combinatorial compression from individual group elements to element-order
counts.  Together they show that the master formula is representation-agnostic
in precisely the claimed way: only the spectra on the chosen representation
change.

The tests are finite and numerical; the all-degree statements (3) and (6)
follow from the preceding eigenvalue and cycle proofs.  Numerical acceptance
requires the direct average to be within $3\times10^{-7}$ of an integer.  The
projector residuals are at machine precision.

\section{Reproducibility}

Run from the repository root:
\begin{verbatim}
PYTHONPATH=for_this_guy/matrix_group_formula_verification \
  .venv/bin/python \
  for_this_guy/matrix_group_formula_verification/\
  representation_variants/verification.py
\end{verbatim}
The marimo notebook allows the maximum degree to be changed interactively and
displays every coefficient comparison and both projector diagnostics.

\end{document}
