\documentclass[11pt]{article}
\usepackage[margin=0.86in]{geometry}
\usepackage{amsmath,amssymb,mathtools,booktabs,microtype,xcolor,fancyhdr,hyperref}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}
\lhead{$B_3\twoheadrightarrow S_3$}\rhead{Proof and verification}\cfoot{\thepage}
\newcommand{\Sgm}{\mathcal S}
\title{\color{navy}\bfseries The Permutation Braid Image $B_3\twoheadrightarrow S_3$\\
\large A non-scalar three-dimensional sigma--MGF test}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle

\begin{abstract}
We construct the adjacent-transposition quotient of $B_3$ in the
three-dimensional permutation representation.  We derive its sigma--MGF
coefficient formula from the three spectral types of $S_3$, and independently
recover the same integers as ranks of explicit Reynolds projectors on tensor
products of symmetric powers.
\end{abstract}

\section{Matrix group and braid relation}
Let
\[
 A=\begin{pmatrix}0&1&0\\1&0&0\\0&0&1\end{pmatrix},\qquad
 B=\begin{pmatrix}1&0&0\\0&0&1\\0&1&0\end{pmatrix}.
\]
These unitary matrices obey $ABA=BAB$.  Sending
$\sigma_1\mapsto A$ and $\sigma_2\mapsto B$ therefore defines a
representation of $B_3$.  Matrix closure gives the six permutation matrices,
so the image is $S_3$ acting on $V=\mathbb C^3$.

\section{Proof and the three-class formula}
For $W_\tau=\bigotimes_a\operatorname{Sym}^{\tau_a}V$, expansion of each
reciprocal determinant followed by Reynolds averaging proves
\[
 [m_\tau]\Sgm_{S_3}=\dim(W_\tau^{S_3})
 =\frac16\sum_{g\in S_3}\prod_a h_{\tau_a}(g),
\]
where $\sum_{r\geq0}h_r(g)t^r=\det(I-tg)^{-1}$.  The three conjugacy
classes have eigenvalues
\[
 (1,1,1),\qquad(1,1,-1),\qquad(1,\omega,\omega^2),
 \quad \omega^3=1.
\]
Their generating series are respectively
\[
 \frac1{(1-t)^3},\qquad \frac1{(1-t)(1-t^2)},\qquad\frac1{1-t^3}.
\]
Consequently, with
\[
 A_r=\binom{r+2}{2},\qquad B_r=\left\lfloor\frac r2\right\rfloor+1,
 \qquad C_r=\mathbf1_{3\mid r},
\]
the proposed class formula becomes the completely explicit identity
\[
 \boxed{[m_\tau]\Sgm_{S_3}=\frac16\left(
 \prod_a A_{\tau_a}+3\prod_a B_{\tau_a}+2\prod_a C_{\tau_a}\right).}
\]
This is also the character-and-power-map formula: Newton's recurrence
$r h_r=\sum_{k=1}^r\operatorname{Tr}(g^k)h_{r-k}$ is obtained by taking
the logarithm of the determinant generating series.

\section{Direct versus formula verification}
The direct route embeds the normalized occupation basis of
$\operatorname{Sym}^rV$ into $V^{\otimes r}$, restricts $g^{\otimes r}$
to that basis, forms the tensor product over the parts of $\tau$, and ranks
the averaged matrix.  The two formula routes compute the power-trace
recurrence over all six matrices and over the three constructed conjugacy
classes.  Thus the direct route does not diagonalize a Reynolds projector by
using its character in advance.

\begin{center}
\begin{tabular}{@{}rrrrr@{}}\toprule
$\tau$&$\dim W_\tau$&projector rank&six elements&three classes\\\midrule
$(1)$&3&1&1&1\\
$(2)$&6&2&2&2\\
$(1,1)$&9&2&2&2\\
$(3)$&10&3&3&3\\
$(2,1)$&18&4&4&4\\
$(2,2)$&36&8&8&8\\\bottomrule
\end{tabular}
\end{center}
All matrices close to exactly six elements, exactly three conjugacy classes
are recovered, and the largest projector error is below $8\times10^{-16}$.

\section{Reproduction}
From the repository root run
\begin{verbatim}
.venv/bin/python for_this_guy/finite_braid_images/
  s3_permutation/verification.py
.venv/bin/marimo edit marimo_notebooks/braid_s3_permutation.py
\end{verbatim}
after joining wrapped lines.  This example verifies the universal finite-image
claim in a non-scalar braid quotient; it is not asserted to be a particular
fusion-space representation of a chosen modular category.
\end{document}

