\documentclass[11pt]{article}
\usepackage[margin=0.84in]{geometry}
\usepackage{amsmath,amssymb,mathtools,booktabs,microtype,xcolor,fancyhdr,hyperref}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}
\lhead{Barnes--Wall family I}\rhead{Extraspecial real Pauli group}\cfoot{\thepage}
\newcommand{\MGF}{\mathfrak M}
\title{\color{navy}\bfseries The Extraspecial Layer of Barnes--Wall Symmetry\\
\large Proof of the four-sector sigma--MGF and direct matrix verification}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle

\begin{abstract}
For the plus-type extraspecial real Pauli group
$E_m\cong 2_+^{1+2m}$ on $V_m=\mathbb C^{2^m}$, we derive the complete
four-sector determinant average.  An independent program constructs every
matrix $\pm X^aZ^b$ and checks 24 one-row and mixed coefficients in
dimensions $2,4,8$.  All comparisons are exact integers and all pass.
\end{abstract}

\section{Representation and coefficient target}
Put $N=2^m$.  On the basis $\{e_x:x\in\mathbb F_2^m\}$ define
\[
 X(a)e_x=e_{x+a},\qquad Z(b)e_x=(-1)^{b\cdot x}e_x.
\]
Then
\[
 E_m=\{\pm X(a)Z(b):a,b\in\mathbb F_2^m\},\qquad |E_m|=2N^2.
\]
For a partition $\tau=(\tau_1,\ldots,\tau_r)$, determinant expansion gives
\begin{equation}\label{eq:target}
 [m_\tau]\MGF_{E_m,m}
 =\frac1{|E_m|}\sum_{g\in E_m}\prod_i
 h_{\tau_i}(\operatorname{spec}g),
\end{equation}
where $\sum_{d\geq0}h_d(\operatorname{spec}g)t^d=
\det(I-tg)^{-1}$.  This is exactly
$\dim(\bigotimes_i\operatorname{Sym}^{\tau_i}V_m)^{E_m}$.

\section{The four spectral sectors}
The central elements $I,-I$ have determinants $(1-t)^N,(1+t)^N$.
For $g=\pm X(a)Z(b)$ one has
\[
 g^2=(-1)^{a\cdot b}I.
\]
Off the center, the trace is zero.  Hence an involution has $N/2$ eigenvalues
of each sign, while an element squaring to $-I$ has $N/2$ eigenvalues each
of $i$ and $-i$.  Their determinants are respectively
$(1-t^2)^{N/2}$ and $(1+t^2)^{N/2}$.

The quadratic form $q(a,b)=a\cdot b$ is plus type.  It has
$(N^2+N)/2$ zeros, including $(0,0)$.  Accounting for both central lifts
therefore gives
\[
 \#\{g\notin Z(E_m):g^2=I\}=N^2+N-2,
 \qquad \#\{g:g^2=-I\}=N^2-N.
\]
The counts sum with the two central elements to $2N^2$.  Averaging proves
\begin{equation}\label{eq:formula}
\boxed{\begin{aligned}
\MGF_{E_m,m}(\mathbf t)=\frac1{2N^2}\Big[&
\prod_j(1-t_j)^{-N}+\prod_j(1+t_j)^{-N}\\
&+(N^2+N-2)\prod_j(1-t_j^2)^{-N/2}\\
&+(N^2-N)\prod_j(1+t_j^2)^{-N/2}\Big].
\end{aligned}}
\end{equation}
Because $-I\in E_m$, the first two terms cancel in odd total degree and the
other two contain only even powers.  Thus every odd-total-degree coefficient
vanishes.

\section{Independent coefficient extraction}
The verifier does not classify matrices after construction.  For each
explicit matrix it computes the eigenvalues, obtains $h_d$ recursively from
the product generating function, forms the integrand in \eqref{eq:target},
and averages.  The formula side separately extracts coefficients from the
four factors in \eqref{eq:formula} using binomial coefficients.
Before coefficient extraction, every matrix is independently checked to have
real entries and to satisfy $g^{\mathsf T}g=I$.

\begin{center}
\begin{tabular}{@{}c@{\quad}c@{\quad}c@{\quad}c@{\qquad}r@{\qquad}r@{}}\toprule
$m$&$N$&$|E_m|$&$\tau$&direct&formula\\\midrule
1&2&8&$(3,1)$&2&2\\
1&2&8&$(2,2)$&3&3\\
2&4&32&$(4)$&5&5\\
2&4&32&$(4,2)$&23&23\\
3&8&128&$(3,1)$&15&15\\
3&8&128&$(6)$&29&29\\
3&8&128&$(4,2)$&190&190\\\bottomrule
\end{tabular}
\end{center}
The full sweep uses eight partitions at each rank, including the odd case
$(1)$ and the mixed cases $(1,1),(3,1),(2,2),(4,2)$.  All 24 pass.

\section{Reproduction}
\begin{verbatim}
.venv/bin/python for_this_guy/barnes_wall_lattice_automorphism_groups/
  extraspecial_real_pauli/verification.py
.venv/bin/marimo edit marimo_notebooks/extraspecial_real_pauli.py
\end{verbatim}
Join each wrapped command onto one shell line.  The computation is exhaustive
at the stated ranks; no random sampling or precomputed character table is
used.
\end{document}
