\documentclass[11pt]{article}
\usepackage[margin=0.82in]{geometry}
\usepackage{amsmath,amssymb,mathtools,booktabs,microtype,xcolor,fancyhdr,hyperref,array}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}\definecolor{soft}{HTML}{F2F6F8}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}
\lhead{Clifford groups}\rhead{Balanced sigma--MGFs}\cfoot{\thepage}
\newcommand{\Sgm}{\mathcal S}\newcommand{\Bmgf}{\mathcal B}
\title{\color{navy}\bfseries Clifford Groups and Balanced Sigma--MGFs\\
\large Proof, phase correction, and explicit one- and two-qubit verification}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle

\begin{abstract}
The ordinary one-sided sigma--MGF does not detect unitary-design order.  We
prove the correct two-sided coefficient statement, isolate a further phase
subtlety, and verify the design-detecting coefficient by closing explicit
projective Clifford matrix groups in dimensions $2$ and $4$.  The computed
frame potentials agree with Haar through degree three and differ at degree
four: $(1,2,5,15)$ versus $(1,2,5,14)$ in dimension two, and
$(1,2,6,29)$ versus $(1,2,6,24)$ in dimension four.
\end{abstract}

\section{One-sided series and the scalar obstruction}
For a finite lift $\widetilde G\leq U(V)$, put
\[
\Sgm_{\widetilde G,V}(\mathbf t)=\frac1{|\widetilde G|}\sum_{g\in\widetilde G}
 \prod_{i\geq1}\det(I-t_i g)^{-1}.
\]
For a compact group $K\leq U(V)$, define instead
\[
\Sgm_{K,V}(\mathbf t)=\int_K\prod_{i\geq1}\det(I-t_i k)^{-1}\,dk
\]
with normalized Haar measure.
Expanding each determinant gives
\[
[m_\tau]\Sgm_{\widetilde G,V}=\dim\left(\bigotimes_j
\operatorname{Sym}^{\tau_j}V\right)^{\widetilde G}.
\]
If the compact group contains the full scalar circle, every positive-degree
one-sided coefficient vanishes.  Hence
\[
\boxed{\Sgm_{K,V}=1.}
\]
for $U(2^n)$ and for the full normalizer $N_{U(2^n)}(P_n)$.  A finite scalar
lift instead obeys only the congruence imposed by its finite scalar center and
may have nonzero lift-dependent coefficients.

\section{The two-sided series, correctly phase neutralized}
For a specified finite lift $\widetilde G\leq U(V)$ define
\[
\Bmgf_{\widetilde G}(\mathbf s,\mathbf t)
=\frac1{|\widetilde G|}\sum_{g\in\widetilde G}
 \prod_i\det(I-s_i g)^{-1}\prod_j\det(I-t_j\bar g)^{-1}.
\]
Character expansion and averaging prove
\begin{equation}\label{eq:two-sided}
[m_\alpha(\mathbf s)m_\beta(\mathbf t)]\Bmgf_{\widetilde G}
=\dim\left(\operatorname{Sym}^{\alpha}V\otimes
\operatorname{Sym}^{\beta}V^*\right)^{\widetilde G}.
\end{equation}

There is a useful refinement.  Replacing a representative $g$ by $zg$
multiplies the coefficient integrand by $z^{|\alpha|-|\beta|}$.  Thus the
entire two-sided series is not intrinsically projective.  Its
\emph{phase-neutral projection}
\begin{equation}\label{eq:neutral}
\Bmgf_G^0=\sum_{|\alpha|=|\beta|}
[m_\alpha m_\beta]\Bmgf_{\widetilde G}\,m_\alpha m_\beta
\end{equation}
is independent of representatives.  Equivalently it is the constant term in
$z$ after $(\mathbf s,\mathbf t)\mapsto(z\mathbf s,z^{-1}\mathbf t)$.
For the full normalizer, continuous scalar integration performs exactly this
projection.  All design coefficients used below lie in \eqref{eq:neutral}.

\section{The design-detecting coefficient}
For $(1^k)=(1,\ldots,1)$, one has
$\operatorname{Sym}^{(1^k)}V=V^{\otimes k}$.  Hence \eqref{eq:two-sided}
gives
\begin{align}
[m_{(1^k)}(\mathbf s)m_{(1^k)}(\mathbf t)]\Bmgf_G^0
&=\dim\operatorname{End}_G(V^{\otimes k})\\
&=\frac1{|G|}\sum_{g\in G}|\operatorname{Tr}g|^{2k}.
\label{eq:frame}
\end{align}
The last quantity is the $k$th unitary frame potential.

Let $P_G^{(k)}$ be the average of
$g^{\otimes k}\otimes\bar g^{\otimes k}$.  It is the orthogonal projection
onto the corresponding $G$-fixed space.  Since $G\subseteq U(d)$, the Haar
fixed space is contained in the $G$-fixed space.  Equality of the coefficient
in \eqref{eq:frame} with the Haar coefficient is equality of the two ranks;
the nested spaces, and therefore their orthogonal projections, are equal.
This proves
\[
\boxed{G\text{ is a unitary }k\text{-design}\iff
\Phi_k(G)=\Phi_k(U(d)).}
\]
Schur--Weyl duality supplies the exact Haar reference
\[
\Phi_k(U(d))=\sum_{\substack{\lambda\vdash k\\\ell(\lambda)\leq d}}
(f^\lambda)^2,
\]
which reduces to $k!$ when $d\geq k$.

\section{Independent matrix verification}
The program constructs
\[
H=2^{-1/2}\begin{pmatrix}1&1\\1&-1\end{pmatrix},\qquad
S=\begin{pmatrix}1&0\\0&i\end{pmatrix}
\]
and, for two qubits, their tensor placements together with CNOT.  Matrices
are multiplied until closure, with equality taken modulo global phase.  The
resulting orders are asserted to be $24$ and $11{,}520$.  No class table or
design theorem is used in the numerical average.  The Haar comparator is
computed independently by the hook-length formula in the Schur--Weyl sum.

\begin{center}
\begin{tabular}{@{}ccrrrr@{}}\toprule
$n$&$d$&$|\overline{\mathrm{Cliff}}_n|$&$k$&direct $\Phi_k$&Haar\\\midrule
1&2&24&1&1&1\\
 &&&2&2&2\\
 &&&3&5&5\\
 &&&4&15&14\\\addlinespace
2&4&11,520&1&1&1\\
 &&&2&2&2\\
 &&&3&6&6\\
 &&&4&29&24\\\bottomrule
\end{tabular}
\end{center}
Thus the first total degree at which the canonical phase-neutral coefficient
can disagree is four, and the frame coefficient does disagree there.
This is precisely the Clifford $3$-design/$4$-design transition established
in \cite{zhu3design,graceful}.
The notebook establishes the displayed frame-potential values only for the
one- and two-qubit projective groups through $k=4$.  The all-qubit 3-design
and failure-of-4-design statements are external theorems supplied by those
references.

As a separate regression test, closing the ordinary one-qubit lift generated
by $H,S$ gives $192$ matrices and scalar center $\mu_8I$.  Its one-sided
coefficients for $\tau=(1),(2),(4)$ vanish, while those for $(8)$ and $(4,4)$
are $1$ and $2$.  This confirms both lift dependence and the scalar selection
rule $8\mid|\tau|$.

\section{Reproduction and scope}
Run
\begin{verbatim}
.venv/bin/python for_this_guy/quantum_matrix_groups/
  clifford_groups/verification.py
.venv/bin/marimo edit marimo_notebooks/clifford_groups.py
\end{verbatim}
after joining each wrapped command onto one line.  The sweep verifies the
multiqubit statement for $n=1,2$ and $k\leq4$; the cited theorems supply the
all-$n$ result.  It makes no inference about odd-prime qudit Clifford groups.

\begin{thebibliography}{9}
\bibitem{zhu3design} H.~Zhu, ``Multiqubit Clifford groups are unitary
3-designs,'' \url{https://arxiv.org/abs/1510.02619}.
\bibitem{graceful} H.~Zhu, R.~Kueng, M.~Grassl, and D.~Gross,
``The Clifford group fails gracefully to be a unitary 4-design,''
\url{https://arxiv.org/abs/1609.08172}.
\end{thebibliography}
\end{document}
