\documentclass[10pt]{article}
\usepackage[margin=0.9in]{geometry}\usepackage{amsmath,amssymb,booktabs,microtype,xcolor,fancyhdr}\usepackage[hidelinks]{hyperref}
\definecolor{navy}{RGB}{24,55,95}\pagestyle{fancy}\fancyhf{}\lhead{$SU(n)$ adjoint}\rhead{Proof and small-$n$ verification}\cfoot{\thepage}\setlength{\headheight}{14pt}
\newcommand{\Sg}{\mathcal S}\newcommand{\CT}{\operatorname{CT}}
\title{\color{navy}\textbf{The adjoint $SU(n)$ sigma-MGF}\\[1mm]\large Proof and exact checks for $n=2,3,4$}
\author{Computational proof companion}\date{17 July 2026}
\begin{document}\maketitle
\begin{abstract}
We derive the compact Molien--Weyl formula for the adjoint representation of
$SU(n)$ and test the manageable cases $n=2,3,4$.  Exact Schur-character
integration confirms the quadratic Killing form, the alternating bracket
tensor, and the symmetric cubic precisely when $n\ge3$.
\end{abstract}

\section{Formula}
The adjoint weights of $SU(n)$ are zero with multiplicity $n-1$ and
$e_i-e_j$ for $i\ne j$.  The general Haar/Cauchy proof yields
\[
 \boxed{\Sg_{SU(n),\mathrm{Ad}}=
 \frac1{n!}\CT_z\left[D_{A_{n-1}}(z)
 \prod_a(1-t_a)^{-(n-1)}
 \prod_{i\ne j}(1-t_a z_i/z_j)^{-1}\right].}
\]
Indeed, expanding the determinant gives
$\sum_\lambda s_\lambda(\mathbf t)\chi_{\mathbb S_\lambda\mathfrak{sl}_n}$;
Haar integration retains exactly the invariant multiplicity.

\section{Low-degree structure}
The Killing form supplies $h_2$, and
$\kappa(X,[Y,Z])$ supplies $e_3$ for every $n\ge2$.  For $n\ge3$ there is also
the symmetric invariant
\[
 (X,Y,Z)\longmapsto\operatorname{Tr}(XYZ+XZY),
\]
which supplies $h_3$.  For $SU(2)$ it vanishes.  Thus, through degree three,
\[
 \Sg_{SU(2),\mathrm{Ad}}=1+h_2+e_3+O_{\ge4},
\]
and, for $n\ge3$,
\[
 \Sg_{SU(n),\mathrm{Ad}}=1+h_2+h_3+e_3+O_{\ge4}.
\]
Chevalley restriction gives
$\Sg(t,0,\ldots)=\prod_{d=2}^n(1-t^d)^{-1}$.

\section{Direct and exact verification}
For each $n=2,3,4$, the notebook generates the $A_{n-1}$ roots, appends the
correct zero weights, and constructs the diagonal adjoint torus matrix.
It compares the determinant with the product over eigenvalues.  It then
computes every partition of degrees zero through three by Jacobi--Trudi and
applies the exact Weyl functional
\[
 \frac1{|W|}\CT(Df)=\CT\left(f\prod_{\alpha>0}(1-z^{-\alpha})\right).
\]

\begin{center}\small
\begin{tabular}{@{}c|c|rrrr@{}}\toprule
$n$&$\dim\mathfrak{sl}_n$&$[h_2]$&$[h_3]$&$[e_3]$&all other $d\le3$\\\midrule
2&3&1&0&1&0\\
3&8&1&1&1&0\\
4&15&1&1&1&0\\\bottomrule
\end{tabular}
\end{center}
All 21 partition-by-case comparisons pass.  The tested representation
dimensions $3,8,15$ are large enough to distinguish the exceptional $n=2$
behavior from the stable $n\ge3$ cubic while remaining directly inspectable.

\paragraph{Reproducibility.} Run \texttt{marimo\_notebooks/lie\_group\_su\_adjoint.py} with marimo.
\end{document}
