\documentclass[10pt]{article}
\usepackage[margin=0.9in]{geometry}\usepackage{amsmath,amssymb,booktabs,microtype,xcolor,fancyhdr}\usepackage[hidelinks]{hyperref}
\definecolor{navy}{RGB}{24,55,95}\pagestyle{fancy}\fancyhf{}\lhead{$E_8$ adjoint}\rhead{Proof and verification strategy}\cfoot{\thepage}\setlength{\headheight}{14pt}
\newcommand{\Sg}{\mathcal S}\newcommand{\CT}{\operatorname{CT}}
\title{\color{navy}\textbf{The compact $E_8$ adjoint sigma-MGF}\\[1mm]\large Proof and dimension-aware verification strategy}
\author{Computational proof companion}\date{17 July 2026}
\begin{document}\maketitle
\begin{abstract}
We prove the exact constant-term formula for the 248-dimensional adjoint
module, verify its root and zero-weight data and its quadratic invariant
exactly, and prove the alternating cubic row algebraically.  We also
record the correct one-copy Chevalley series and explain why it does not
determine the multivariate sigma-MGF.
\end{abstract}

\section{Exact formula}
For every unitary compact-group representation,
\[
 \Sg_{K,V}=\int_K\prod_a\det(I-t_a\rho(g))^{-1}dg
 =\sum_\lambda\dim(\mathbb S_\lambda V)^K s_\lambda(\mathbf t).
\]
Equivalently,
$[m_\tau]\Sg_{K,V}=\dim(\bigotimes_i\operatorname{Sym}^{\tau_i}V)^K$.
The adjoint $E_8$ weights are the 240 roots and zero with multiplicity eight.
Weyl integration therefore proves
\[
 \boxed{\Sg_{E_8,248}=\frac1{696729600}\CT_z\left[D_{E_8}(z)
 \prod_a(1-t_a)^{-8}\prod_{\alpha\in\Phi(E_8)}
 (1-t_a z^\alpha)^{-1}\right].}
\]

\section{Universal adjoint tensors}
For a compact simple Lie algebra, the Killing form $\kappa$ is a unique
symmetric quadratic invariant.  In degree three,
\[
 c(x,y,z)=\kappa(x,[y,z])
\]
is invariant, nonzero, and fully alternating.  Moreover
\[
(\bigwedge^3\mathfrak g)^G\cong
\operatorname{Hom}_G(\bigwedge^2\mathfrak g,\mathfrak g),
\]
and for the simple Lie algebra $\mathfrak e_8$ the bracket spans this
Hom-space.  Chevalley restriction gives basic invariant degrees
$2,8,12,14,18,20,24,30$, so $(\operatorname{Sym}^3\mathfrak g)^G=0$.
Since the only invariant in $\mathfrak g^{\otimes3}$ is alternating, the
Schur type $(2,1)$ also has multiplicity zero.  Consequently
\[
 \boxed{\Sg_{E_8,248}=1+h_2+e_3+O_{\ge4}.}
\]
Chevalley restriction gives the one-copy series
\[
 \boxed{\Sg(t,0,\ldots)=\prod_{d\in\{2,8,12,14,18,20,24,30\}}
 (1-t^d)^{-1}.}
\]
This specialization cannot see $e_3$ and therefore does not determine the
full sigma-MGF.
Thus the degree-at-most-three monomial signature is
\[
([m_{(2)}],[m_{(1,1)}],[m_{(3)}],[m_{(2,1)}],[m_{(1,1,1)}])
=(1,1,0,0,1).
\]

\section{Executed quadratic check and optional sparse cubic confirmation}
The notebook generates the $E_8$ root orbit (240 vectors), appends eight zero
weights, and constructs a $248$ by $248$ diagonal unitary torus matrix.  It
checks the determinant integrand directly and integrates the quadratic Schur
characters exactly with the Weyl-denominator identity.
\begin{center}\small
\begin{tabular}{@{}c|rrrr@{}}\toprule
$\lambda$&$\varnothing$&$(1)$&$(2)$&$(1,1)$\\\midrule
proposed&1&0&1&0\\
exact Weyl check&1&0&1&0\\\bottomrule
\end{tabular}
\end{center}
The computed dimension is 248, with 240 roots and eight zero weights; all
quadratic entries pass.  Floating-point work is confined to the illustrative
torus matrix, while invariant multiplicities use rational weights and exact
integer sums.

The cubic multiplicities
$(\operatorname{Sym}^3,\mathbb S_{(2,1)},\bigwedge^3)=(0,0,1)$ are therefore
proved algebraically above.  For an optional mechanical confirmation,
construct only the zero-weight part of
$\bigwedge^3\mathfrak e_8$ and impose the eight simple-root raising operators.
The expected nullspace is one-dimensional and represented by $c(x,y,z)$; the
analogous expected nullities on $\operatorname{Sym}^3$ and
$\mathbb S_{(2,1)}$ are zero.  This sparse route avoids materializing spaces
of dimensions on the order of $\binom{248}{3}$.  It is not reported as an
executed computation in this artifact.

\paragraph{Reproducibility.} Run \texttt{marimo\_notebooks/lie\_group\_e8.py} with marimo.
\end{document}
