\documentclass[11pt]{article}
\usepackage[margin=0.84in]{geometry}
\usepackage{amsmath,amssymb,mathtools,booktabs,microtype,xcolor,fancyhdr,hyperref}
\definecolor{navy}{HTML}{17365D}\definecolor{teal}{HTML}{147D78}
\hypersetup{colorlinks=true,linkcolor=navy,urlcolor=teal}
\pagestyle{fancy}\setlength{\headheight}{14pt}\fancyhf{}
\lhead{Rooted-tree automorphisms}\rhead{Exact plethystic recursion}\cfoot{\thepage}
\newcommand{\MGF}{\mathcal M}
\title{\color{navy}\bfseries Rooted-Tree Automorphism Groups\\
\large Wreath-product proof, exact matrix tests, and the limit boundary}
\author{Computational verification note}\date{17 July 2026}
\begin{document}\maketitle

\begin{abstract}
We prove the exact recursive sigma--MGF law for automorphisms of any finite
rooted tree.  Explicit leaf-permutation matrix groups for complete binary
trees through height three and ternary trees through height two give 30 exact
mixed-coefficient comparisons.  The document separates this proved finite
recursion from the still-open general random-tree $\Lambda$-limit.
\end{abstract}

\section{Wreath-product lemma}
Let a finite permutation group $H$ act on $X$, and let $H\wr S_m$ act on
$X\times[m]$.  Define
\[
\MGF_H(\mathbf t)=\frac1{|H|}\sum_{h\in H}
\prod_i\det(I-t_i h)^{-1}.
\]
Write an element of the wreath product as $(h_1,\ldots,h_m;\sigma)$.  On a
top cycle $c=(i_1\cdots i_\ell)$, one circuit accumulates
$h_c=h_{i_\ell}\cdots h_{i_1}$ and block elimination gives
\[
\det(I-tg|_c)=\det(I-t^\ell h_c).
\]
Uniform independent $h_i$ make the products $h_c$ uniform and independent
over the top cycles.  Averaging over $S_m$ and using its cycle index proves
\begin{equation}\label{eq:wreath}
\boxed{\sum_{m\ge0}u^m\MGF_{H\wr S_m}(\mathbf t)=
\exp\left(\sum_{\ell\ge1}\frac{u^\ell}{\ell}
\MGF_H(t_1^\ell,t_2^\ell,\ldots)\right).}
\end{equation}

\section{Arbitrary rooted trees}
If the root branches have isomorphism types $A$ with multiplicities $m_A$,
then internal automorphisms and permutations of equal branches give
\[
\operatorname{Aut}(T)\cong\prod_A(\operatorname{Aut}(A)\wr S_{m_A}).
\]
The leaf representation is the corresponding direct sum.  Determinants
multiply on direct sums, so \eqref{eq:wreath} yields
\begin{equation}\label{eq:tree}
\boxed{\MGF_T(\mathbf t)=\prod_A[u_A^{m_A}]
\exp\left(\sum_{\ell\ge1}\frac{u_A^\ell}{\ell}
\MGF_A(\mathbf t^\ell)\right).}
\end{equation}
For the complete $b$-ary tree, $G_h=G_{h-1}\wr S_b$, starting with the
trivial action on one leaf.

\section{Exact computational verification}
The verifier recursively enumerates every wreath-product permutation and
materializes its zero-one matrix.  Distinct matrix keys and the identity
\[
|G_h|=b!\,|G_{h-1}|^b
\]
independently check each closure.  For a matrix $M$, the direct symmetric
characters are obtained over exact rationals from Newton's identity
\[
d h_d(M)=\sum_{r=1}^d\operatorname{Tr}(M^r)h_{d-r}(M).
\]
The other side is computed from the exact cycle-index average in
\eqref{eq:wreath}, with truncated multivariate polynomial arithmetic.  Thus
the two routes share neither matrices nor coefficient extraction.

\begin{center}
\begin{tabular}{@{}rrrrr@{}}\toprule
$b$&height&leaves&base order&group order\\\midrule
2&1&2&1&2\\
2&2&4&2&8\\
2&3&8&8&128\\
3&1&3&1&6\\
3&2&9&6&1296\\\bottomrule
\end{tabular}
\end{center}
For every row the program compares coefficients indexed by
$(1),(2),(3),(1,1),(2,1),(2,2)$; all 30 comparisons pass exactly.

\section{What is and is not proved}
Equation \eqref{eq:tree} is an exact theorem for each finite tree.  For a
Galton--Watson tree it becomes a distributional recursion after conditioning
on branch-type counts.  Also,
\[
\log|\operatorname{Aut}(T)|=\sum_i\log|\operatorname{Aut}(T_i)|
+\sum_A\log(m_A!)
\]
is an additive tree functional; Gaussian-type asymptotics for this scalar
quantity are known.  Neither fact alone proves convergence of the full random
series $\MGF_T(\mathbf t)$.  A general limiting rooted-tree
$\Lambda$-distribution therefore remains open, and the verifier makes no such
claim.

\section{Reproduction and reference}
\begin{verbatim}
.venv/bin/python for_this_guy/limiting_lambda_claims/
  rooted_tree_automorphisms/verification.py
.venv/bin/marimo edit marimo_notebooks/rooted_tree_automorphisms.py
\end{verbatim}
Join wrapped commands onto one line.  For automorphism-count asymptotics, see
M.~Olsson and S.~Wagner, \emph{The distribution of the number of
automorphisms of random trees},
\url{https://arxiv.org/abs/2211.12801}.
\end{document}
