\documentclass[11pt]{article}
\usepackage[margin=0.82in]{geometry}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs,tabularx,array}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=true,linkcolor=NavyBlue,urlcolor=BrickRed,citecolor=NavyBlue]{hyperref}
\usepackage{microtype}
\usepackage{fancyhdr}
\usepackage{enumitem}

\definecolor{ink}{HTML}{17212B}
\definecolor{accent}{HTML}{315B7D}
\definecolor{pale}{HTML}{EDF4F8}
\definecolor{passgreen}{HTML}{16794A}
\color{ink}
\pagestyle{fancy}
\fancyhf{}
\lhead{Tensor-product local unitary groups}
\rhead{Proof and verification}
\cfoot{\thepage}
\setlength{\headheight}{14pt}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}
\newcommand{\Sym}{\operatorname{Sym}}
\newcommand{\End}{\operatorname{End}}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Exp}{\operatorname{Exp}_{\sigma}}
\newcommand{\SMG}{\operatorname{SMG}}
\newcommand{\PASS}{\textcolor{passgreen}{\textbf{PASS}}}
\newcommand{\summarybox}[1]{\begin{center}\fcolorbox{accent}{pale}{\parbox{0.91\linewidth}{#1}}\end{center}}

\title{\textbf{Balanced Sigma-MGFs for Tensor-Product Local Unitary Groups}\\[3pt]
\large Proof, exact formulas, and direct matrix verification}
\author{Reproducibility note}
\date{17 July 2026}

\begin{document}
\maketitle

\summarybox{\textbf{Outcome.} For
$G=\prod_{a=1}^kU(d_a)$ on $V=\bigotimes_{a=1}^k\mathbb C^{d_a}$,
the ordinary one-sided sigma-MGF is exactly $1$. The nontrivial object is the
balanced two-sided series involving both $V$ and $V^*$. Its stable power-sum
coefficient is $z_\lambda^{k-2}$, its bipartite single-state coefficient is a
restricted partition number, and its multilinear coefficient is an exact
product of Schur--Weyl centralizer dimensions. A reproducible marimo notebook
constructs the Kronecker-product matrices and compares direct Haar estimates
with these formulas. All 46 numerical comparisons, 4 construction checks, and
5 scalar-center witnesses pass.}

\section{Representation and the correction at the outset}

Put $V_a=\mathbb C^{d_a}$ and
\[
 G=U(d_1)\times\cdots\times U(d_k),\qquad
 V=V_1\otimes\cdots\otimes V_k.
\]
The representation is the concrete matrix homomorphism
\begin{equation}
 \rho(g_1,\ldots,g_k)=g_1\otimes\cdots\otimes g_k
 \in U(D),\qquad D=\prod_{a=1}^kd_a.
\end{equation}
For a partition $\tau=(\tau_1,\ldots,\tau_\ell)$, write
\[
 \Sym^\tau V=\bigotimes_{j=1}^{\ell}\Sym^{\tau_j}V,
 \qquad |\tau|=\sum_j\tau_j.
\]
The ordinary coefficient is
\[
 [m_\tau]\SMG_{G,V}=\dim(\Sym^\tau V)^G.
\]

\textbf{Proposition 1 (one-sided collapse).}
\begin{equation}
 \boxed{\SMG_{G,V}=1.}
\end{equation}

\textit{Proof.}
For $z\in U(1)$ the group contains
$(zI_{d_1},I_{d_2},\ldots,I_{d_k})$, whose matrix in (1) is $zI_V$.
It acts on $\Sym^\tau V$ as $z^{|\tau|}I$. If $|\tau|>0$, one can choose
$z$ with $z^{|\tau|}\ne1$, so a fixed vector must be zero. Only degree zero
remains. \hfill$\square$

The same center explains why local-unitary entanglement invariants use a state
$\psi$ and its conjugate $\bar\psi$: the weights $z$ and $z^{-1}$ cancel only
when the two total degrees agree.

\section{The balanced series and Molien formula}

Introduce independent alphabets $\mathbf s$ and $\mathbf t$ and define
\begin{equation}
 \mathcal B_{G,V}(\mathbf s,\mathbf t)
 =\int_G \Exp\!\left(X_V(g)h_1(\mathbf s)\right)
          \Exp\!\left(X_{V^*}(g)h_1(\mathbf t)\right)\,dg.
\end{equation}
Coefficient extraction gives
\begin{align}
 \mathcal B_{G,V}
 &=\sum_{\alpha,\beta}b_{\alpha,\beta}
          m_\alpha(\mathbf s)m_\beta(\mathbf t),\\
 b_{\alpha,\beta}
 &=\dim\left(\Sym^\alpha V\otimes\Sym^\beta V^*\right)^G
  =\dim\Hom_G(\Sym^\beta V,\Sym^\alpha V).
\end{align}
The center now gives the balanced selection rule
\begin{equation}
 \boxed{b_{\alpha,\beta}=0\quad\text{unless}\quad
 |\alpha|=|\beta|.}
\end{equation}

For finitely many variables, (3) is the balanced Molien integral
\begin{equation}
 \mathcal B_{G,V}
 =\int_G\prod_i\det(I-s_i\rho(g))^{-1}
         \prod_j\det(I-t_j\rho(g)^{-1})^{-1}\,dg.
\end{equation}
If $x^{(a)}_1,\ldots,x^{(a)}_{d_a}$ are the eigenvalues of $g_a$, then the
eigenvalues of $\rho(g)$ are $x^{(1)}_{i_1}\cdots x^{(k)}_{i_k}$.
Consequently (7), together with the Weyl integration formula for each local
factor, is an exact constant-term prescription in every dimension.

\section{Stable power-sum formula}

For every positive integer $r$, Kronecker products satisfy the directly
testable identity
\begin{equation}
 \Tr(\rho(g)^r)=\prod_{a=1}^k\Tr(g_a^r).
\end{equation}
Thus, for a partition $\lambda$,
$p_\lambda(\rho(g))=\prod_a p_\lambda(g_a)$, and product Haar measure gives
\begin{equation}
 \mathbb E_G\!\left[p_\lambda(\rho(g))
 \overline{p_\mu(\rho(g))}\right]
 =\prod_{a=1}^k\mathbb E_{U(d_a)}
 \left[p_\lambda(g_a)\overline{p_\mu(g_a)}\right].
\end{equation}
Let $z_\lambda=\prod_i i^{m_i(\lambda)}m_i(\lambda)!$. Schur orthogonality,
or equivalently the stable Hall inner product, gives
\begin{equation}
 \mathbb E_{U(d)}[p_\lambda(U)\overline{p_\mu(U)}]
 =\delta_{\lambda\mu}z_\lambda
 \quad\text{when }d\ge |\lambda|,|\mu|.
\end{equation}
Combining (9) and (10), if $N\le\min_a d_a$ then
\begin{equation}
 \mathbb E_G[p_\lambda(\rho)\overline{p_\mu(\rho)}]
 =\delta_{\lambda\mu}z_\lambda^k
 \quad (|\lambda|,|\mu|\le N).
\end{equation}
Each determinant exponential contributes a factor $p_\lambda/z_\lambda$.
Therefore through bidegree $(N,N)$,
\begin{equation}
 \boxed{\mathcal B_{G,V}\equiv
 \sum_{|\lambda|\le N}z_\lambda^{k-2}
 p_\lambda(\mathbf s)p_\lambda(\mathbf t).}
\end{equation}
For $k=1,2,3$, the coefficients are respectively
$z_\lambda^{-1},1,z_\lambda$.

\section{Exact single-state coefficients}

For partitions $\lambda^{(1)},\ldots,\lambda^{(k)}\vdash m$, define the
generalized Kronecker coefficient
\[
 g(\lambda^{(1)},\ldots,\lambda^{(k)})
 =\dim\left([\lambda^{(1)}]\otimes\cdots\otimes[\lambda^{(k)}]\right)^{S_m}.
\]
Applying Schur--Weyl duality to each $V_a^{\otimes m}$ and then taking
diagonal $S_m$-invariants proves
\begin{equation}
 \Sym^m(V_1\otimes\cdots\otimes V_k)
 \cong\bigoplus_{\lambda^{(1)},\ldots,\lambda^{(k)}\vdash m}
 g(\lambda^{(1)},\ldots,\lambda^{(k)})
 \bigotimes_{a=1}^kS_{\lambda^{(a)}}V_a,
\end{equation}
where terms with $\ell(\lambda^{(a)})>d_a$ vanish. Schur orthogonality now
gives the all-dimensional identity
\begin{equation}
 \boxed{b_{(m),(m)}=
 \sum_{\substack{\lambda^{(a)}\vdash m\\
                  \ell(\lambda^{(a)})\le d_a}}
 g(\lambda^{(1)},\ldots,\lambda^{(k)})^2.}
\end{equation}

For $k=2$, $g(\lambda,\mu)=\delta_{\lambda\mu}$. Hence
\begin{equation}
 \boxed{b_{(m),(m)}=
 \#\{\lambda\vdash m:\ell(\lambda)\le\min(d_1,d_2)\}.}
\end{equation}
This is $p(m)$ when both dimensions are at least $m$. In stable range, setting
each alphabet in (12) to one variable yields the equivalent formula
\begin{equation}
 \boxed{b_{(m),(m)}=\sum_{\lambda\vdash m}z_\lambda^{k-2}.}
\end{equation}

For completeness, general $b_{\alpha,\beta}$ also has an exact finite formula.
Decompose each $\Sym^{\alpha_j}V$ by (13), then combine local Schur functors
with Littlewood--Richardson coefficients. If the resulting multiplicity of
$\bigotimes_aS_{\mu^{(a)}}V_a$ is $M_\alpha(\mu^{(1)},\ldots,\mu^{(k)})$,
then
\begin{equation}
 b_{\alpha,\beta}=\sum_{\substack{\mu^{(1)},\ldots,\mu^{(k)}\\
                          \ell(\mu^{(a)})\le d_a}}
 M_\alpha(\mu^{(1)},\ldots,\mu^{(k)})
 M_\beta(\mu^{(1)},\ldots,\mu^{(k)}).
\end{equation}
This separates the two combinatorial ingredients: generalized Kronecker
coefficients enter the symmetric powers of a tensor product, while
Littlewood--Richardson coefficients enter products of symmetric powers.

\section{Exact multilinear coefficients}

For $\alpha=\beta=(1^r)$, $\Sym^{(1^r)}V=V^{\otimes r}$, so
\begin{equation}
 b_{(1^r),(1^r)}=\dim\End_G(V^{\otimes r}).
\end{equation}
Because the local factors act independently,
\[
 \End_G(V^{\otimes r})\cong
 \bigotimes_{a=1}^k\End_{U(d_a)}(V_a^{\otimes r}).
\]
If $f^\lambda$ is the number of standard Young tableaux of shape $\lambda$,
Schur--Weyl duality and the hook-length formula give
\begin{equation}
 \boxed{b_{(1^r),(1^r)}=
 \prod_{a=1}^k\sum_{\substack{\lambda\vdash r\\
                    \ell(\lambda)\le d_a}}(f^\lambda)^2.}
\end{equation}
When every $d_a\ge r$, the inner sum is $|S_r|=r!$, and therefore
\begin{equation}
 \boxed{b_{(1^r),(1^r)}=(r!)^k.}
\end{equation}
This is the dimension of the stable family of contraction diagrams indexed by
$k$ independent permutations in $S_r$.

\section{Verification design}

The accompanying implementation uses two independent routes.

\textbf{Exact route.}
Integer partitions are generated recursively. The hook-length formula computes
$f^\lambda$, which is substituted into (19). Equation (15) is evaluated by
restricted partition counting, and (16) by direct computation of $z_\lambda$.
No sampled matrices enter these tables.

\textbf{Direct matrix route.}
For every local factor, a complex Gaussian matrix $Z$ is QR-factorized and the
diagonal phases of $R$ are removed. The resulting $Q$ is Haar distributed on
$U(d_a)$. The code explicitly forms
$\rho=Q_1\otimes\cdots\otimes Q_k$, checks $\rho^*\rho=I$, and evaluates its
power traces. Newton's recurrence
\begin{equation}
 mh_m(\rho)=\sum_{j=1}^m\Tr(\rho^j)h_{m-j}(\rho)
\end{equation}
then produces characters of $\Sym^mV$ directly. The sampled target quantities
are
\begin{align*}
 &\mathbb E\prod_jh_{\tau_j}(\rho) &&\text{(ordinary one-sided)},\\
 &\mathbb E[p_\lambda(\rho)\overline{p_\mu(\rho)}]
     &&\text{(balanced power-sum)},\\
 &\mathbb E|h_m(\rho)|^2 &&\text{(single-state)},\\
 &\mathbb E|\Tr\rho|^{2r} &&\text{(multilinear)}.
\end{align*}
The last two are character inner products, hence exactly the invariant
multiplicities in (14) and (19). Numerical integration is corroboration, not a
replacement for the proofs above.

\clearpage
\section{Results}

Each case used 6,000 deterministic-seed Haar samples. A numerical row passes
when the absolute error is at most six empirical standard errors plus a small
fixed rounding margin. Matrix unitarity and trace factorization were also
checked at tolerance $10^{-11}$.

\begin{center}
\begin{tabularx}{\linewidth}{@{}Xlrrrr@{}}
\toprule
Representation & Target & Direct estimate & Formula & Std. error & Result\\
\midrule
$U(3)$ & $|h_3|^2$ & 1.0367 & 1 & 0.0253 & \PASS\\
$U(3)$ & $|\Tr\rho|^4$ & 2.0318 & 2 & 0.0593 & \PASS\\
$U(2)\times U(3)$ & $|h_2|^2$ & 1.9533 & 2 & 0.0546 & \PASS\\
$U(2)\times U(3)$ & $|p_{(2)}|^2$ & 3.8911 & 4 & 0.0638 & \PASS\\
$U(2)\times U(2)$ & $|h_3|^2$ & 2.0797 & 2 & 0.1219 & \PASS\\
$U(2)^3$ & $|h_2|^2$ & 4.0815 & 4 & 0.1641 & \PASS\\
$U(2)^3$ & $|\Tr\rho|^4$ & 8.2364 & 8 & 0.6379 & \PASS\\
$U(2)^3$ & $p_{(2)}\overline{p_{(1,1)}}$ & $0.0032$ & 0 & 0.0284 & \PASS\\
\bottomrule
\end{tabularx}
\end{center}

The exact calculations make the stable-range boundary visible:
\begin{center}
\begin{tabular}{@{}llrrl@{}}
\toprule
Dimensions & Sector & Degree & Exact & Stable expression\\
\midrule
$(2,2)$ & single-state & 3 & 2 & $p(3)=3$ (not applicable)\\
$(2,2)$ & multilinear & 3 & 25 & $(3!)^2=36$ (not applicable)\\
$(2,3)$ & multilinear & 3 & 30 & $(3!)^2=36$ (not applicable)\\
$(2,2,2)$ & multilinear & 3 & 125 & $(3!)^3=216$ (not applicable)\\
\bottomrule
\end{tabular}
\end{center}
Thus the dimension restrictions are not cosmetic: applying a stable formula
outside $r,m\le\min_a d_a$ produces wrong answers.

The scalar-center test separately constructs, for each degree $1\le n\le5$,
$z=e^{2\pi i/(n+1)}$ and the group matrix
$\rho(zI_2,I_3)=zI_6$. Its action on degree $n$ is $z^n\ne1$ in every case,
giving five direct matrix witnesses for Proposition~1.

\section{Reproduction and scope}

The folder \texttt{for\_this\_guy/local\_unitary\_tensor\_products} contains:
\begin{itemize}[leftmargin=1.5em]
 \item \texttt{marimo_notebooks/local_unitary_tensor_products.py}, the interactive marimo report;
 \item \texttt{verification.py}, the direct and exact implementation;
 \item \texttt{test\_verification.py}, deterministic regression tests;
 \item this \LaTeX{} source and its compiled PDF.
\end{itemize}
The notebook exposes the sample count and reports every tested coefficient, not
only the representative rows printed here. Its global $U(3)$ case is explicitly
labeled as a comparator; it is not conflated with the local tensor-product
groups. All local-unitary cases remain grouped in this single family because
they are specializations of the same representation (1) and the same proof.

\end{document}
